\LARGE A double-sampling extension of the German National Forest Inventory for design-based small area regression estimation of timber volume resources on forest district levels \\ \vspace{2ex} \large Large scale application to the federal state of Rhineland-Palatinate \vspace{-1ex}

### 2018-02-13 Draft_Form/tab/modacc_modterms.tex
update 13_02_2018
update 13_02_2018

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Sun Jan 07 17:11:26 2018
\begin{table*}[!htbp]
	\begin{center}
	\caption{Accuracy metrics for the two OLS regression models on the cluster level. Interaction terms are indicated by ':'. () give the respective values on the plot level.} 
	\label{tab:modacc_modterms}
    {\small %
	\begin{tabular}{llccc}
  \hline
model terms & model & $R^2_{adj}$ & RMSE & RMSE\% \\ 
  \hline \hline
meanheight + stddev + meanheight$^2$ +  & full model &  0.59 & 101.62  & 33.57 \\
treespecies + ALSyear + & & (0.48) &  (139.22) & (45.98) \\ 
meanheight:treespecies + \\ meanheight:ALSyear + meanheight:stddev + \\ stddev:ALSyear &&& \\ \\

meanheight + stddev + meanheight$^2$ + & reduced model  & 0.56  & 106.08 & 35.4 \\
ALSyear + meanheight:ALSyear + & & (0.45) & (144.13) & (47.60) \\
meanheight:stddev + stddev:ALSyear &&& \\ \\
\hline
\hline
\end{tabular}
}%
\end{center}
\end{table*}
%\endgroup

>>>>

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Sun Jan 07 17:11:26 2018
\begin{table*}[!htbp]
	\begin{center}
	\caption{Accuracy metrics for the two OLS regression models on the cluster level. Interaction terms are indicated by ':'. () give the respective values on the plot level.} 
	\label{tab:modacc_modterms}
    {\small %
	\begin{tabular}{llccc}
  \hline
model terms & model & $R^2_{adj}$ & RMSE & RMSE\% \\ 
  \hline \hline
meanheight + stddev + meanheight$^2$ +  & full model &  0.58 & 90.11  & 29.76 \\
treespecies + ALSyear + & & (0.48) &  (139.22) & (45.98) \\ 
meanheight:treespecies + \\ meanheight:ALSyear + meanheight:stddev + \\ stddev:ALSyear &&& \\ \\

meanheight + stddev + meanheight$^2$ + & reduced model  & 0.55  & 95.23 & 31.65 \\
ALSyear + meanheight:ALSyear + & & (0.45) & (144.13) & (47.60) \\
meanheight:stddev + stddev:ALSyear &&& \\ \\
\hline
\hline
\end{tabular}
}%
\end{center}
\end{table*}
%\endgroup

### 2018-02-13 Draft_Form/tex/Appendices.tex
update 13_02_2018
update 13_02_2018

%\appendix
%\addcontentsline{toc}{section}{Appendices}
%\section*{Appendices}
%\section{First appendix}
%\section{Second appendix}

\appendix
\section{Appendix}

% ----------------------------------------------------------------------- %
% A.1
% ----------------------------------------------------------------------- %
%\subsection{Model accuracies for $CHM$ variables}

%\input{tab/mod_chm_table.tex}\newpage

>>>>

\appendix
\section{Appendix}
\label{sec:appendix_saerlp}

% -------------------------------------- %
\subsection*{R-squared on cluster level}

The R$^2$ on the cluster level is calculated using the number of plots $M(x)$ of each cluster in order to weight for the varying number of plots on which $Y_c(x)$ and $\hat{Y}_{c}(x)$ are based on.

\begin{equation*}\label{eq:R_squared_clust}
R^2 =  \frac{\sum_{x \in s_2} \Big(\frac{M(x)}{\bar{M_2}}\Big)^2 \Big(\hat{Y}_{c}(x)-\hat{\bar{Y}}_c\Big)^2}{\sum_{x \in s_2} \Big(\frac{M(x)}{\bar{M_2}}\Big)^2 \Big(Y_{c}(x)-\hat{\bar{Y}}_c\Big)^2}
\end{equation*}

\noindent $Y_c(x)$ and $\hat{Y}_{c}(x)$ are the predicted and observed local densities on the cluster level calculated according to equations \ref{eq:locdens_clust} and \ref{eq:globresids}. $\hat{\bar{Y}}_c$ is the estimated sample mean corresponding to the weighted mean over all observed local densities on the cluster level (equ. \ref{eq:srs_est_clusters}).\par

% -------------------------------------- %
\subsection*{RMSE on cluster level}

The same weights $M(x)$ are also applied to calculate the RMSE on the cluster level. $n_2$ is the number of clusters used in the modeling frame.

\begin{equation*}\label{eq:RMSE_clust}
RMSE= \sqrt{\frac{1}{n_{2}}  \sum_{x \in s_2} \Big(\frac{M(x)}{\bar{M}_2}\Big)^2 \Big(\hat{Y}_{c}(x)-Y_c(x)\Big)^2}
\end{equation*}

\noindent The \textit{relative} or \textit{normalized} RMSE is calculated by setting the absolute RMSE in relation to the estimated sample mean $\hat{\bar{Y}}_c$:

\begin{equation*}\label{eq:NRMSE_clust}
RMSE[\%] = \frac{RMSE}{\hat{\bar{Y}}_c}
\end{equation*}

\noindent Generally note that the weights $\frac{M(x)}{\bar{M_2}} \equiv 1$ if the number of plots per cluster is constant.

### 2018-02-13 Draft_Form/tex/Discussion.tex
update 13_02_2018
update 13_02_2018

The aim of this study was to investigate the performance of design-based estimators for small area estimation of mean standing timber volume on two spatial forest management levels in Germany. It was of particular interest to gather information about the estimation error levels that can be realized using the German NFI data, characterized by low sampling intensities in the area of interests, as terrestrial phase information in the two-phase estimators. To address these research questions, we applied the simple random sampling (SRS), the pseudo-small (\psmall{}) and the extended pseudo-synthetic (\extpsynth{}) estimator to 45 and 405 small area units in the German state of Rhineland-Palatinate.\par

@@

For several FR units, it was observed that the \psmall{} and the \extpsynth{} estimator can produce larger variances than the SRS estimator. It is important to note that this is in perfect agreement with the theory of both two-phase estimators and can theoretically appear if the residual variance in the small area, which generally constitutes the dominating part of the two-phase variance, turns out to be much higher than the variance of the terrestrial data in the small area. The empirical findings of our study suggest that such cases can particularly occur if moderate or poor model fits within a small area are combined with small terrestrial sample sizes ($\leq 5$) in the small area. A closer look on these small areas thus might reveal the reason for the poor prediction performance and help to improve the model fit. Nonetheless, it should be kept in mind that small terrestrial sample sizes can also cause the SRS estimator to not reflect the actual variation of the local density within a small area. In this case, the two-phase variance estimate might be larger but more realistic. Whereas a visual analysis of aerial images, remote sensing data or stand maps might give some further evidence for or against this hypothesis, a definite proof is practically infeasible.\par

We were also able to empirically confirm that the \extpsynth{} estimator generally produces slightly smaller variances and estimation errors than the \psmall{}. This is most probably caused by marginally smaller model residuals due to the intercept adjustment to the terrestrial data in the small area unit, which is primarily a means to ensure the zero-mean-residual property of the \extpsynth{}. However, our analysis indicated that the difference between the two estimators are negligible for sample sizes $\geq$ 10 due to their asymptotic relationship. Furthermore, one or more clusters not entirely included in the small area unit did reassuringly not have a notable impact on the estimates of \extpsynth{}. This empirical evidence can however not be generalized to any application of \extpsynth{} under cluster sampling. It thus seems recommendable to calculate both \psmall{} and \extpsynth{}, and subsequently compare their results. If no suspicious deviations occur, we consider the \extpsynth{} as the estimator of choice.\par

>>>>

The aim of this study was to investigate the performance of design-based estimators for small area estimation of mean standing timber volume on two spatial forest management levels in Germany. It was of particular interest to gather information about the estimation error levels that can be realized using the German NFI data, characterized by low sampling intensities in the area of interests, as terrestrial phase information in the two-phase estimators. To address these research questions, we applied the simple random sampling (SRS), the pseudo-small (\psmall{}) and the extended pseudo-synthetic (\extpsynth{}) estimator for cluster sampling to 45 and 405 small area units in the German state of Rhineland-Palatinate.\par

@@

For several FR units, it was observed that the \psmall{} and the \extpsynth{} estimator can occasionally produce larger variances than the SRS estimator. It is important to note that this is in perfect agreement with the theory of both two-phase estimators and can theoretically appear if the residual variance in the small area, which generally constitutes the dominating part of the two-phase variance, turns out to be much higher than the variance of the terrestrial data in the small area. The empirical findings of our study suggest that such cases can particularly occur if moderate or poor model fits within a small area are combined with small terrestrial sample sizes ($\leq 5$) in the small area. A closer look on these small areas thus might reveal the reason for the poor prediction performance and help to improve the model fit. Nonetheless, it should be kept in mind that small terrestrial sample sizes can also cause the SRS estimator to not reflect the actual variation of the local density within a small area. In this case, the two-phase variance estimate might be larger but more realistic. Whereas a visual analysis of aerial images, remote sensing data or stand maps might give some further evidence for or against this hypothesis, a definite proof is practically infeasible.\par

We were also able to empirically confirm that the \extpsynth{} estimator generally produces slightly smaller variances and estimation errors than the \psmall{}. This is most probably caused by marginally smaller model residuals due to the intercept adjustment to the terrestrial data in the small area unit, which is primarily a means to ensure the zero-mean-residual property of the \extpsynth{}. However, our analysis indicated that the difference between the two estimators are negligible for sample sizes $\geq$ 10 due to their asymptotic equivalency. Furthermore, one or more clusters not entirely included in the small area unit did reassuringly not have a notable impact on the estimates of \extpsynth{}. This empirical evidence can however not be generalized to any application of \extpsynth{} under cluster sampling. It thus seems recommendable to calculate both \psmall{} and \extpsynth{}, and subsequently compare their results. If no suspicious deviations occur, we consider the \extpsynth{} as the estimator of choice.\par

### 2018-02-13 Draft_Form/tex/Materials_and_Methods(2).tex
update 13_02_2018
update 13_02_2018

where $nha_i$ is the number of trees per hectare represented by tree $i$. The local densities on cluster level can then be used to derive the estimated spatial mean $\hat{Y}_c$ and its estimated variance $\hat{\var}(\hat{Y}_c)$ for any given spatial unit for which $n_2 \geq 2$ ($n_2$ denoting the number of sample units, i.e. clusters):

\begin{subequations}\label{eq:srs_est_cluster}
	\begin{align}
\hat{Y}_c &= \frac{\sum_{x \in s_2} M(x) Y_{c}(x)}{\sum_{x \in s_2} M(x)} \\
\hat{\var}(\hat{Y}_c) &= \frac{1}{n_2(n_2 - 1)}\sum_{x \in s_2}\Big(\frac{M(x)}{\bar{M_2}}\Big)^2(Y_{c}(x) - \hat{Y}_c)^2
	\end{align}
\end{subequations}

@@

\noindent with $\hat{R}_{c}=Y_{c}(x)-\pmb{Z}_{c}^{\top}(x)\hat{\pmb{\beta}}_{c,s_2}$ being the empirical model residuals at the cluster level, which by construction of OLS satisfy the important \textit{zero mean residual property}, i.e. $\frac{\sum_{x \in s_{2}} M(x) \hat{R}_{c}(x)}{\sum_{x \in s_{2}} M(x)}=0$.\\

@@

\begin{subequations}\label{eq:compsG}
	\begin{align}
	\hat{\bar{\pmb{Z}}}_{c,G} &= \frac{\sum_{x \in s_{1,G}} M_{G}(x) \pmb{Z}_{c,G}(x)}{\sum_{x \in s_{1,G}} M_{G}(x)} \label{eq:meanZcG} \hspace{3ex} &\text{where} \hspace{3ex}
	\pmb{Z}_{c,G}(x) =  \frac{\sum_{l=1}^{L} I_{G}(x_l)\pmb{Z}(x_l)}{M_{G}(x)}\\		
	Y_{c,G}(x) &= \frac{\sum_{l=1}^{L} I_{G}(x_l)\pmb{Y}(x_l)}{M_{G}(x)}  &\text{and} \hspace{3ex}
	\hat{Y}_{c,G}(x)=\hat{\bar{\pmb{Z}}}_{c,G}^{\top}\hat{\pmb{\beta}}_{c,s_2}\\
	\bar{\hat{R}}_{2,G} &= \frac{\sum_{x \in s_{2,G}} M_{G}(x) \hat{R}_{c,G}(x)}{\sum_{x \in s_{2,G}} M_{G}(x)} \hspace{3ex} &\text{where} \hspace{3ex}
	\hat{R}_{c,G}(x) = Y_{c,G}(x)-\hat{Y}_{c,G}(x)
	\end{align}
\end{subequations}

@@

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} derived from the ALS data as well as the \treespecies{} variable derived from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{}, \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.59 and RMSE of 101.61 \mha{} and 33.6\%). A stratification to the ALS acquisition years substantially improved the model accuracy and thus proved to be an effective means in accounting for the artificially introduced noise in the data set due to ALS quality variations and time-gaps between the ALS and the terrestrial survey. However, the stratification led to a highly unbalanced dataset when stratifying according to the \treespecies{} variable. For this reason, a species individual modeling within each \alsyear{} stratum remained infeasible, but might have further improved the model accuracy. An additional evaluation of the model residuals within each ALS acquisition year stratum revealed that the model accuracies substantially varied between the strata (table \ref{tab:adj_r2_within}). Values above the overall \adjrsq{} were particularly achieved in ALS acquisition years close to the terrestrial survey date (0.60 to 0.65 on the cluster level). This effect was substantially diminished when dropping the \alsyear{} variable from the model term.\par

>>>>

where $nha_i$ is the number of trees per hectare represented by tree $i$. The local densities on cluster level can then be used to derive the estimated spatial mean $\hat{\bar{Y}}_c$ and its estimated variance $\hat{\var}(\hat{Y}_c)$ for any given spatial unit for which $n_2 \geq 2$ ($n_2$ denoting the number of sample units, i.e. clusters):

\begin{subequations}\label{eq:srs_est_clusters}
	\begin{align}
\hat{\bar{Y}}_c &= \frac{\sum_{x \in s_2} M(x) Y_{c}(x)}{\sum_{x \in s_2} M(x)} \\
\hat{\var}(\hat{\bar{Y}}_c) &= \frac{1}{n_2(n_2 - 1)}\sum_{x \in s_2}\Big(\frac{M(x)}{\bar{M_2}}\Big)^2(Y_{c}(x) - \hat{\bar{Y}}_c)^2
	\end{align}
\end{subequations}

@@

with

\begin{equation}\label{eq:globresids}
\hat{R}_{c}=Y_{c}(x)-\pmb{Z}_{c}^{\top}(x)\hat{\pmb{\beta}}_{c,s_2} = Y_{c}(x) - \hat{Y}_{c}(x)
\end{equation}

\noindent being the empirical model residuals at the cluster level, which by construction of OLS satisfy the important \textit{zero mean residual property}, i.e. $\frac{\sum_{x \in s_{2}} M(x) \hat{R}_{c}(x)}{\sum_{x \in s_{2}} M(x)}=0$.\\

@@

\begin{subequations}\label{eq:compsG}
	\begin{align}
	\hat{\bar{\pmb{Z}}}_{c,G} &= \frac{\sum_{x \in s_{1,G}} M_{G}(x) \pmb{Z}_{c,G}(x)}{\sum_{x \in s_{1,G}} M_{G}(x)} \label{eq:meanZcG} \hspace{3ex} &\text{where} \hspace{3ex}
	\pmb{Z}_{c,G}(x) =  \frac{\sum_{l=1}^{L} I_{G}(x_l)\pmb{Z}(x_l)}{M_{G}(x)}\\		
	Y_{c,G}(x) &= \frac{\sum_{l=1}^{L} I_{G}(x_l)Y(x_l)}{M_{G}(x)}  &\text{and} \hspace{3ex}
	\hat{Y}_{c,G}(x)=\hat{\bar{\pmb{Z}}}_{c,G}^{\top}\hat{\pmb{\beta}}_{c,s_2}\\
	\bar{\hat{R}}_{2,G} &= \frac{\sum_{x \in s_{2,G}} M_{G}(x) \hat{R}_{c,G}(x)}{\sum_{x \in s_{2,G}} M_{G}(x)} \hspace{3ex} &\text{where} \hspace{3ex}
	\hat{R}_{c,G}(x) = Y_{c,G}(x)-\hat{Y}_{c,G}(x)
	\end{align}
\end{subequations}

@@

The model with highest \adjrsq{} and lowest RMSE was achieved with the variables \meanheight{}, \meanheight{}$^2$, \stddev{} and \alsyear{} derived from the ALS data as well as the \treespecies{} variable derived from the classification map as main effect terms. Additionally, interaction terms between \meanheight{} and \alsyear{}, \stddev{} and \alsyear{}, \meanheight{} and \stddev{}, and \meanheight{} and \treespecies{} were included. The model yielded an \adjrsq{} of 0.48 and an RMSE of 140.62 \mha{} (46.7\%) on the plot level (table \ref{tab:modacc_modterms}, full model). While the study of \citet{hill2017a} only provide the model evaluations on the plot level, the two-phase estimators described in section \ref{sec:SAestimators} derive and apply the regression coefficients and the residuals on the aggregated cluster level. We thus re-evaluated the model as used in the estimators also on the cluster level (formulas given in Appendix) and found the model accuracies to be higher than on the plot level (\adjrsq{} of 0.59 and RMSE of 101.61 \mha{} and 33.6\%). A stratification to the ALS acquisition years substantially improved the model accuracy and thus proved to be an effective means in accounting for the artificially introduced noise in the data set due to ALS quality variations and time-gaps between the ALS and the terrestrial survey. However, the stratification led to a highly unbalanced dataset when stratifying according to the \treespecies{} variable. For this reason, a species individual modeling within each \alsyear{} stratum remained infeasible, but might have further improved the model accuracy. An additional evaluation of the model residuals within each ALS acquisition year stratum revealed that the model accuracies substantially varied between the strata (table \ref{tab:adj_r2_within}). Values above the overall \adjrsq{} were particularly achieved in ALS acquisition years close to the terrestrial survey date (0.60 to 0.65 on the cluster level). This effect was substantially diminished when dropping the \alsyear{} variable from the model term.\par
